---
title: "R Notebook"
output: html_notebook
---

```{r}
library(gtools)
library(tidyverse)
source("day_1/prob.R")

titanic<- read_csv("day_1/5_quiz_and_homework/data/titanic.csv")

titanic %>% 
 distinct(class)
```
```{r}

titanic_new <- titanic %>% 
  mutate(pass_or_crew =
    ifelse(
      class %in% 
        c("engineering crew", "victualling crew", "restaurant staff", "deck crew"), 
      "crew", "passenger")) 

```

```{r}
titanic_new %>% 
tabyl(pass_or_crew, survived) %>% 
  adorn_totals(where = c("row", "col"))
```
p(A | B) = p(A given B) = p(A AND B) / P(B)

p(survived | crew) = p(survived & crew) / p(crew)
                   = (211/890) / (890/ 2207) = 0.5879018

```{r}
# (211/890) / (890/ 2207)
# 
# (500/1317) / (1317/2207)
```

p(survived | passenger) = p(survived & passenger) / p(passenger)
                        = (500/1317) / (1317/2207) = 0.6362104
                        
                        
1.2 

Question 1:

Probability of rolling a six sided die once and getting a number less than five is 4/6 (2/3 or 0.666)

Question 2:

```{r}
s_three_dice <- rolldie(3, nsides = 6)
#therea re 216 possible outcomes rolling a three 6-sided dice at once.
```
Question 3:

In an experiment where we roll four dice:

how many outcomes are in the event where the total of the four dice equals 5
, 6 or 7)?

what is the probability of this event?

what assumptions do we apply in answering part ii?

```{r}

four_die <- rolldie(4, nsides = 6)

four_die %>% 
  filter(X1 + X2 + X3 + X4 > 4 & X1 + X2 + X3 + X4 < 8)

# There are 34 possible outcomes. 

34/1296
#the probability is 0.02623457

# we assume that we can get repeated values for each die ie multiple 1's or 2's
```

Question 4

```{r}
permutations(
  n = 5,
  r = 5, 
  v = c(1, 2, 3, 4, 5),
  repeats.allowed = FALSE
) %>% 
  as.tibble() %>% 
  filter(V1 == 5, V2 == 4, V3 == 3, V4 == 2, V5 == 1)

#1/120 chance that they sit in order of decreasing height, left to right.
```

```{r}

```























