---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(modelr)
```
```{r}
wine_quality_red <- read_csv("Weekend Homework/wine_quality_red.csv")
wine_quality_white <- read_csv("Weekend Homework/wine_quality_white.csv")

wine_quality_red 
wine_quality_white
```

Traits which make a good wine:
acidity - high acidity, sharper the taste of the wine, feel lighter bodied
tannin - phenolic compounds in wine, bitter taste, higher in red wines
alcohol - higher alcohol bolder taste, low alcohol wines - feel lighter
sweetness - refers to levels of residual sugar (sweet wines have higher residual sugar)
- all traits must be balanced.

variables:
outcome = quality

vif - various inflation factor

input =

1 - fixed acidity: most acids involved with wine are fixed or nonvolatile (do not evaporate readily)

2 - volatile acidity: the amount of acetic acid in wine, hight levels can cause vinegar taste

3 - citric acid: small amount in wines, can add ‘freshness’ and flavour

4 - residual sugar: the amount of sugar residual after fermentation, wines with less than 1 gram/liter and  greater than 45 grams/liter are sweet

5 - chlorides: "salts" in the wine

6 - free sulfur dioxide: the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial growth and the oxidation of wine

7 - total sulfur dioxide: total free and bound forms of S02; in low concentrations undetectable but free SO2 concentrations over 50 ppm, can affect smell and taste

8 - density: the density of wine is close to that of water depending on the percent alcohol and sugar content

9 - pH: describes how acidic or basic a wine is, most wines are between 3-4 on the pH scale

10 - sulphates: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant

11 - alcohol: the percent alcohol content of the wine

4 groups of variables:


```{r}

wine_quality_red_type <- wine_quality_red %>% 
  mutate(wine_id = str_c("r", wine_id, sep = "_"),
         colour = "red", 
         .after = wine_id
         )

wine_quality_white_type <- wine_quality_white %>% 
  mutate(wine_id = str_c("w", wine_id, sep = "_"),
         colour = "white", 
         .after = wine_id
         )

```

```{r}
wine_quality_combined <- rbind(wine_quality_red_type, wine_quality_white_type) %>% 
  select(-wine_id)
```

```{r}
wine_quality_combined

head(wine_quality_combined)

summary(wine_quality_combined) # no NA's in the data...
```
```{r}
# data for both red and white wines follows a relatively normal distribution
wine_quality_combined %>% 
  ggplot() +
  aes(x = quality) +
  geom_histogram(colour = "white", bins = 30) +
  facet_wrap(~ colour) +
  scale_x_continuous(limits = c(0, 10))


```
split into numerical and categorical variables:

outcome: quality, 
categorical: colour, region

numerical: fixed_acidity, volatile_acidity, citric_acid, p_h,

residual_sugar, chlorides,
free_sulfur_dioxide, total_sulfur_dioxide, density,  sulphates, alcohol


```{r}
#converted categoricals to factors
wine_quality_combined <- wine_quality_combined %>% 
  mutate(colour = as.factor(colour),
         region = as.factor(region))





ggpairs(wine_quality_combined, progress =  FALSE)
```

```{r}
# numerical variables(acidity) correlations with quality
wine_quality_combined %>% 
  select(quality, fixed_acidity, volatile_acidity, citric_acid, p_h,) %>% 
  ggpairs(progress = FALSE)

# none of the variables relating to acidity of the wine have a particularly high crrelation with the wine quality. fixed and volatile acidity seem to cause a reduction in the quality of the wine(weak and relatively weak), citric acid level has a very weak positive correlation, pH also has a very weak/no correlation with the quality of the wine. 
```
```{r}
# other numerical variables(not-acidity) correlations with quality

wine_quality_combined %>% 
  select(-c(fixed_acidity, volatile_acidity, citric_acid, p_h, colour, region)) %>% 
  ggpairs(progress = FALSE) +
  theme_grey(base_size = 1) +# this bit just changes the axis label font size so we can see
  theme(strip.text = element_text(size = 5)) 

# similar to other numerical variables none of the variables here have a particuaraly high correlation
# with the wine quality. 
# the highest correlations (low/moderate) are density, chlorides, and alcohol level
# other variables  sulphates, sulphur dioxide levels and, residual sugar, sulphates

```

```{r}
# categorical variables

wine_quality_combined %>% 
  select(quality, colour, region) %>% 
  ggpairs(progress = FALSE)

# from ggpairs plots neither categorical variable, colour or region, appears to have a strong correlation
# witht the quality score of the wine. 

```

From ggpairs plots the top candidate variables to explore were (in order):

volatile_acidity
density
alcohol
chlorides

```{r}
alias(quality ~ ., data = wine_quality_combined)
```
# first predictor

```{r}
model1a <- lm(quality ~ volatile_acidity, data = wine_quality_combined)
model1b <- lm(quality ~ density, data = wine_quality_combined)
model1c <- lm(quality ~ alcohol, data = wine_quality_combined)
model1d <- lm(quality ~ chlorides, data = wine_quality_combined)

# model 1a - quality ~ volatile acidity
autoplot(model1a)
# residuals look ok
# relatively normally distributed
# not too much heteroscedacisity
summary(model1a)
# r2 value is relatively low, 0.06349, but the effect of volatile_acidity is significant.
```

```{r}
# model 1a - quality ~ volatile acidity
autoplot(model1b)
# residuals are not distributed well as outlier appears to alter residuals away from 0 
# data does appear to be normally distributed
# large amount of heteroscedasticity observed...
# based upon this would probably discount density as a variable for our linear model.

summary(model1b)
# r2 value, low although density has a signifcant effect on quality... 
```
```{r}
quality ~ alcohol
autoplot(model1c)
# residuals well distiributed.
# data does appear to be normally distributed
# no heteroscedasicity observed

summary(model1c)
# r2 value low, 0.1832, (higher than volatility) and effect is significant

```
```{r}
# model1d quality ~ chlorides

autoplot(model1d)
# residuals are not distributed well as outlier appears to alter residuals away from 0 
# data does appear to be normally distributed
# large amount of heteroscedasticity observed...
# based upon this would probably discount chlorides as a variable for our linear model.

summary(model1d)
# r2 low, effect signifcant but does not meet criteria for linear model 
```
best model so far is quality ~ alcohol
model1c <- lm(quality ~ alcohol, data = wine_quality_combined)




# second predictor

```{r}
wine_resid <- wine_quality_combined %>% 
  add_residuals(model1c, var = "resid") %>% 
  select(resid, volatile_acidity, density, chlorides)


wine_resid %>% 
  ggpairs(progress = FALSE)

#both volatile_acidity and density have the highest correlation with the residuals
```

```{r}
model2a <- lm(quality ~ alcohol + volatile_acidity, data = wine_quality_combined)

model2b <- lm(quality ~ alcohol + density, data = wine_quality_combined)

model2c <- lm(quality ~ alcohol + chlorides, data = wine_quality_combined)

```

```{r}
# model2a = quality ~ alcohol + volatile_acidity 

 autoplot(model2a)

# residual distribution looks good, 
# data relatively normally distributed
# low heteroscedasticity

summary(model2a)

# adj r2 low, 0.2387,  effect significant 


```
```{r}
# model2b <- quality ~ alcohol + density

autoplot(model2b)
# residual distribution looks good, 
# data relatively normally distributed
# low heteroscedasticity

summary(model2b)
# adj r2 = 0.183 but effect not significant, would not use density in the model
```
```{r}
# modoel2c = quality ~ alcohol +chlorides 

autoplot(model2c)
# residual distribution looks ok, bit skewed
# data relatively normally distributed
# low heteroscedasticity

summary(model2c)
# adj r2 = 0.1904
# chlorides does signifcantly alter the quality. 
```

model2a with quality ~ alcohol + volatile acidity has the highest adjusted r2

```{r}
anova(model1c, model2a)
# low p value tells us the model is signficant so we should include the volatile acidity
# variable
```


# third predictor

```{r}
wine_resid2 <- wine_quality_combined %>% 
  add_residuals(model2a, var = "resid") %>% 
  select(resid, density, chlorides)

wine_resid2 %>% 
  ggpairs(progress = FALSE)
```


```{r}
model3a <- lm(quality ~ alcohol + volatile_acidity + chlorides, data = wine_quality_combined)
model3b <- lm(quality ~ alcohol + volatile_acidity + density, data = wine_quality_combined)
```

```{r}
# model3a <- lm(quality ~ alcohol + volatile_acidity + chlorides, data = wine_quality_combined)

autoplot(model3a)
# residuals look consisten,
# data normally distributed
# low heteroscedasticity

summary(model3a)
# adj r2 is 0.2386 but effect not signifcant

```
```{r}
# model3b <- lm(quality ~ alcohol + volatile_acidity + density, data = wine_quality_combined)

autoplot(model3b)
# residuals look consistent,
# data normally distributed
# low heteroscedasticity

summary(model3b)
# adj r2 is 0.2457 , effect of density is signifcant

```
```{r}
anova(model2a, model3b)
```

# remaining predictors 

```{r}
wine_resid_3 <- wine_quality_combined %>% 
  add_residuals(model3b, var = "resid") %>% 
  select(-c(quality, alcohol, volatile_acidity, density))

wine_resid_3 %>% 
  ggpairs(progress = FALSE) +
  theme_grey(base_size = 1) +# this bit just changes the axis label font size so we can see
  theme(strip.text = element_text(size = 5)) 

# none of the predictors that were excluded in the beginning correlate well with the 
# residuals of the final model.
  
```

## Model Interpretation

The quality rating of wine is affected by the alcohol concentration, volatile_acidity + density.

for a wine with density alcohol % of 4, volatile_acidity of 0.7 and density of 0.98,

quality = -35.51606 + 0.38 * alcohol + -1.49647 * volatile_acidity + 37.98219 * density
quality = -35.51606 + 4 * 0.38 + -1.49647 * 0.7 + 37.98219 * 0.98
quality = 2.178957

Note: from homework answers it is important to remove variables that are highly correlated with each other. In this instance it would have been important to drop density as this is highly correlated with alcohol content.




