---
title: "Week2_Day1_Homework"
author: "DC"
date: "2023-03-27"
output: html_document
---
##Q1

```{r}
library(tidyverse)
library(tidyr)
library(janitor)
card_base <- read.csv("data/CardBase.csv")
customer_base <- read.csv("data/CustomerBase.csv")
fraud_base <- read.csv("data/FraudBase.csv")
transaction_base <- read.csv("data/TransactionBase.csv")
hat_observations <- read.csv("data/hat_observations.csv")
```

```{r}
card_base <- clean_names(card_base)
customer_base <- clean_names(customer_base)
fraud_base <- clean_names(fraud_base)
transaction_base <- clean_names(transaction_base)
```

##Q2

```{r}
card_base
customer_base
```
```{r}
inner_join(card_base, customer_base, "cust_id")
```
Answer:
left_join(card_base, customer_base, by = "cust_id")
right_join(customer_base, card_base, by = "cust_id")

##Q3


```{r}
full_join(fraud_base, transaction_base, "transaction_id")
```
Answer:
The resulting row number is 10,000 (same as transaction details) - hence, all records in fraud table have a match in transaction details (can check this by checking where NAs are - all in the fraud_flag column)

##Q4

```{r}
#unique(transaction_base)
```

```{r}
card_base
transaction_base
```

```{r}
semi_join(card_base, transaction_base, by = c("card_number" = "credit_card_id"))
```




##Q5

```{r}
hat_observations
```

```{r}
(hat_observations_new <- hat_observations %>% 
  separate(observation, c("hat_colour", "hat_type")))
              
```

Answer:

```{r}
# hat_observations <- read_csv("data/hat_observations.csv")

hat_observations <- hat_observations %>% 
  separate(observation, sep = ",", into = c("hat_colour", "hat_type"))

hat_observations
```




##Q6

```{r}
hat_observations_new %>% 
  unite(date, c("day", "month", "year"), sep = "/") %>% 
  filter(hat_type == "beret") %>% 
  arrange(desc(observation_count)) %>% 
  head(1) %>% 
  pull(date)
  
  
```
Other method:

hat_observations_date <- hat_observations %>% 
  unite("date", day:year, sep = "-") 

hat_observations_date %>% 
  filter(hat_type == "beret") %>% 
  group_by(date) %>% 
  summarise(total = sum(observation_count)) %>% 
  slice_max(total)
  
  
#Extension

##Q1

```{r}
(card_customer_base <- inner_join(card_base, customer_base, "cust_id"))

(card_customer_base_new <- card_customer_base %>% 
    relocate(card_number, .after = customer_vintage_group))

```

```{r}
(fraud_transaction_base <- full_join(fraud_base, transaction_base, 
                                     "transaction_id"))

(fraud_transaction_base_new <- fraud_transaction_base %>% 
  relocate(credit_card_id, .before = transaction_id) %>% 
    relocate(fraud_flag, .after = transaction_segment))
```


```{r}
(joined_dataset <- inner_join(card_customer_base_new, fraud_transaction_base_new, 
                            c("card_number" = "credit_card_id"))) %>% 
  filter(!is.na(fraud_flag))

(final_joined_dataset <- joined_dataset %>% 
  filter(!is.na(fraud_flag)) %>% 
  relocate(card_number, .before = card_family))

```
```{r}
colSums(is.na(joined_dataset))
```

Answer:

```{r}
card_base %>% 
  inner_join(customer_base, by = "cust_id") %>% 
  inner_join(transaction_base, by = c("card_number" = "credit_card_id")) %>% 
  inner_join(fraud_base, by = "transaction_id")
```




## Q2

```{r}
(exam_scores <- read.csv("data/exam_scores.csv"))
```

```{r}
exam_scores %>% 
  pivot_longer(cols = starts_with("exam"),
               names_to = "exam_question",
               values_to = "exam_scores") %>% 
  separate(exam_question, c("exam_", "question"), sep = "_") %>% 
  select(id, question, exam_scores)
  
```

Answer removing the Q:

exam_scores <- read_csv("data/exam_scores.csv")

exam_scores_long <- exam_scores %>% 
  pivot_longer(exam_Q1:exam_Q10, names_to = "exam_question", values_to = "score") 

exam_scores_long %>% 
  separate(exam_question, sep = "Q", into = c("extra_column", "exam_question")) %>% 
  select(-extra_column)













