---
title: "R Notebook"
output: html_notebook
---

```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(janitor)

library(tidyverse)
titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

# MVP

## Q1

```{r}
titanic_set
```
```{r}
clean_titanic_set <- titanic_set %>% 
 filter(!is.na(survived)) %>% 
  mutate(survival = factor(survived, levels = c(0,1), labels = c("No", "Yes")), 
         sex = as.factor(sex),
         age_status = as.factor(ifelse(age <= 16, "child", "adult")),
         class = factor(pclass, levels = c(3,2,1), labels = c("Low", "Med", "High")),
         embarked = as.factor(embarked)
         ) %>% 
   na.omit() %>% 
  select(sex, age_status, class, embarked, sib_sp, parch, survival) 
 
  
```



# Question 2

```{r}
library(GGally)
clean_titanic_set %>% 
  ggpairs(progress = FALSE)
# from ggpairs plot we can see that sex, age_status, class, embarkment all influence survivial
```
```{r}
clean_titanic_set %>% 
  ggplot() +
  aes(x = survival, fill = sex) %>% 
  geom_bar(position = "fill")

clean_titanic_set %>% 
  ggplot() +
  aes(x = survival, fill = class) %>% 
  geom_bar(position = "fill")

clean_titanic_set %>% 
  ggplot() +
  aes(x = survival, fill = age_status) %>% 
  geom_bar(position = "fill")

```


## Question 3

```{r}


n_data <- nrow(clean_titanic_set)

test_index <- sample(1:n_data, size = n_data*0.2)

titanic_test  <- slice(clean_titanic_set, test_index)

titanic_train <- slice(clean_titanic_set, -test_index)

titanic_test %>% 
   tabyl(survival)

titanic_train %>% 
  tabyl(survival)

# the data is not perfectly balanced but is close enough based on the notes.
```
## Question 4
```{r}
titanic_fit <- rpart(
  formula = survival ~ ., 
  data = titanic_train, 
  method = 'class'
)

rpart.plot(titanic_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4)

rpart.plot(titanic_fit, 
           yesno = 2, 
           fallen.leaves = TRUE, 
           faclen = 6, 
           digits = 4, 
           type = 4, 
           extra = 101)

```
```{r}
rpart.rules(titanic_fit, cover = TRUE)
```


## Question 5

The decision tree tells us that based on our data:
- the root node, tells us when all the data is examined together the most likely outcome is survival on the titanic.
- the probability of surviving base on the root node is 0.6735.
- the biggest predictor of survival is sex not being male and if you are female the probability of survival is 0.9557 with 48.33% of the data points pass through this node.
- if you are male you have a 0.4079 chance of survival with 51.79% of the data passing through this leaf node, 
- if you are male the next biggest predictor of your survival is if you have greater than 1 siblings or spouses on board the titanic. If you have no sibling or spouse you have a 0.3265 chance of survival being yes,
- if you do not have a sibling or spouse aboard your chnace of survival is 0.5556 with 18.37% of the data passing through this leaf node.
- the next predictor of survival would be where you embarked from. ifyou embarked at Queensland or South Hampton, you had a 0.5 chance of survival with 13.61% of the data being passed to this node. If you did not embark at Queensland or Sputh Hampton your chance of survival was 0.7143 percent with 4.76% of the data passing through this node.
- the final determinant of survival in our decision tree is whether you travelling with a child or parents, or whether you had a nanny aboard. If you had nanny aboard your chance of survival was 0.667 with 6.12% of the data passing through this node. Whether you were travelling with parents or children your chance of survival was 0.3636 with 7.48% of the data passing through this leaf node. 

Females have a highest overall chance or surviving.
Males with no siblings or spouses aboard had the lowest chance of surviving

## Question 6

```{r}
library(modelr)

titanic_test_pred <- titanic_test %>%
  add_predictions(titanic_fit, type = 'class')

library(caret)

confusionMatrix(titanic_test_pred$pred, titanic_test_pred$survival)
```
For this data set the confusion matrix  tells us:
there were 7 TNs, 8FPs, 5FNs, 16 TPs.
63.89% of the data was classified correctly
CIs are (0.4622 to 0.7918)
sensitivity indicates the model is good, not great, at identifying postives instances
specificity is 0.6667, indicates the model is very good at identifying negative instances.

## Question 7

```{r}
library(ranger)

rf_classifier <- ranger(survival ~ ., 
                        data = titanic_train, 
                        importance = "impurity", 
                        num.trees = 1000, 
                        mtry = 2, 
                        min.node.size = 5)

rf_classifier

importance(rf_classifier)


```

sex is the most important variable, followed by sib_sp, age_status, parch, and class. 

```{r}
titanic_test_pred <- titanic_test %>%
  mutate(pred = predict(rf_classifier, data = titanic_test)$predictions)

confusionMatrix(titanic_test_pred$pred, titanic_test_pred$survival)
```








