---
title: "Week 5 Day 2 Homework"
output: html_notebook
---

Question 1.

Load in the libraries you need, and look at the backpack data. 
Take note of variable names and dimensions. 

```{r}
library(tidyverse)
backpack <- CodeClanData::backpack

dim(backpack)
glimpse(backpack)
```

Question 2.

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
backpack %>% 
  ggplot() +
  geom_density(aes(x = backpack_weight)) +
  labs(x = "Backpack Weight", y = "Count Number of Backpacks",
       title = "Distribution of Backpack Weight",
       subtitle = "from all students") +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))

# The distribtuion of the backpack weight is shown with most backpacks, weighing around 10,
# there are also some weighing less than 10, some weighing between 10 and 20 and very few weighing
# more than 20kg or 30kg. either density or histogram plot show the data distribution.

backpack %>% 
  ggplot() +
  geom_histogram(aes(x = backpack_weight), fill = "powderblue", colour = "black")  +
  theme_bw() +
  labs(x = "Backpack Weight", y = "Count Number of Backpacks",
       title = "Distribution of Backpack Weight",
       subtitle = "from all students") +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))

```

Question 3.

Are there differences in backpack weight between sex?

```{r}
#The graph below shows that the total backpack weight in males and females is different. 
backpack %>% 
  ggplot() +
  geom_col(aes(x = sex, y = backpack_weight, fill = sex)) +
  labs(x = "Sex", y = "Total Backpack Weight",
       title = "Backpack Weight",
       subtitle = "in Males and Females") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))

#There is a difference in the total weight of the backpacks in both makes and females. 
#Females are carrying a total weight of more than males. 


```


Question 4.

What is the relationship between body weight and backpack weight?

```{r}
#The graph below shows a scatter graph with body weight plotted and backpack weight plotted on each axis.
# The line shows that there is a very weak linear relationship between increasing backpack weight and body weight.
# It may suggest that a higher backpack weight correlates with higher backpack weight.
backpack %>% 
  ggplot() +
  aes(x = backpack_weight, y = body_weight) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Backpack Weight", y = "Body Weight",
       title = "Backpack Weight and Body Weight",
       subtitle = "in all Students") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))
```


Question 5.

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.

```{r}
#The graph below shows a scatter graph with body weight plotted and backpack weight plotted on each axis for females and males.
# In females the line shows that there is a very weak linear relationship between increasing backpack weight and body weight.
# It may suggest that a higher backpack weight correlates with higher backpack weight.
# The same is true for the male students.
backpack %>% 
  ggplot() +
  aes(x = backpack_weight, y = body_weight, colour = sex) +
  geom_point() +
  facet_wrap(~ sex) +
  geom_smooth(method = "lm") +
  labs(x = "Backpack Weight", y = "Body Weight",
       title = "Relationship between Backpack Weight and Body Weight",
       subtitle = "in Male and Female Students") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))

```


Question 6.

What proportion of students report back pain?
```{r}
# 68 of 100 students report back pain and 32 out of 100 do not. The graph below shows
# the number of students with reporting with back pain as true vs the number of students without as false.
backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>% 
  ggplot() +
  aes(x = back_problems, fill = back_problems) +
  geom_bar() +
  labs(x = "Back Pain", y = "Number of Students",
       title = "Number of Students with Back Pain",
       subtitle = "TRUE or False for all Students") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8)) +
  ylim(0, 80) 

```
Question 7.

Is this proportion different between men and women?

```{r}
#The proportion differs between men and women as from the graph below around 37 men report
#no backpain and a small number (less than 10) report back pain as TRUE. For women,
# around 31 report back pain as false and about 24 report back pain as true. So although
# there are more women in the study (55f vs 45 M) around half of emales report backpain as true.
# \
# backpack %>% 
#   mutate(male_female = str_count(sex, "Male")) %>% 
#   summarise(sum(male_female))
table(backpack)

backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>%
  # group_by(back_problems) %>% 
  # summarise(sum(back_problems == TRUE)) %>%
  ggplot() + 
  aes(x = back_problems, fill = sex) +
  geom_bar() +
  facet_wrap(~ sex) +
  labs(x = "Back Pain", y = "Number of Students",
       title = "Number of Students with Back Pain",
       subtitle = "True or False for all Students") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8)) +
  ylim(0, 40) 

```

Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}
backpack %>% 
  mutate(back_problems = as.logical(back_problems)) %>%
  ggplot() + 
  aes(x = back_problems, y = backpack_weight, fill = sex) +
  geom_col() +
  facet_wrap(~ sex) +
  labs(x = "Back Pain", y = "Backpack Weight",
       title = "Backweight and Back Pain",
       subtitle = "True or False for all Students") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8)) +
  ylim(0, 40) 

## The graph below shows that in both males and females a higher total backpack weight is present in those
##male and female students suffering with back pain.
```

```{r}
backpack %>% 
  ggplot() +
  aes(x = sex, y = back_problems, fill = sex) +
  geom_col() +
  labs(x = "Sex", y = "Number of Students with Back Pain",
       title = "Sex and Back Pain") +
  theme_bw() +
  theme(plot.title.position = "plot",
        plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 8))

# The graph below shows that a much higher number of female students have back pain than males.
```

```{r}
backpack %>%
  mutate(back_problems = as.logical(back_problems)) %>%
  ggplot() +
  aes(x = backpack_weight,  y = ratio, size = back_problems, colour = back_problems) +
  geom_point() +
  geom_smooth(method = "lm", size = 0.5) +
  labs(x = "Backpack Weight", y = "Body Weight to Backpack Weight Ratio",
       title = "Relationship between Backpack Weight and Body Weight to Backpack Weight Ratio") +
  theme_bw() +
  theme(plot.title.position = "plot",
         plot.title = element_text(size = 14),
         plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
         axis.text = element_text(size = 8))

#The graph below shows that in students with or without backpack pain there is a linear relationship between
#bodyweight to backpack weight ratio and backpack weight.
```

```{r}
backpack %>%
  ggplot() +
  aes(x = backpack_weight,  y = ratio, colour = sex) +
  geom_point() +
  geom_smooth(method = "lm", size = 0.5) +
  labs(x = "Backpack Weight", y = "Backpack Weight to Body Weight Ratio",
       title = "Relationship between Backpack Weight and Body Weight to Backpack Weight Ratio",
       subtitle ="in Males and Females") +
  theme_bw() +
  theme(plot.title.position = "plot",
         plot.title = element_text(size = 14),
         plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
         axis.text = element_text(size = 8))

# The graph below shows that in both males and females backpack weight to body weight ratio correlates linearly with 
# backpack weight.
```
```{r}
backpack %>%
  mutate(back_problems = as.logical(back_problems)) %>%
  ggplot() +
  aes(x = backpack_weight,  y = ratio) +
  geom_point() +
  geom_smooth(method = "lm", size = 0.5) +
  facet_grid(sex ~ back_problems) +
  labs(x = "Backpack Weight", y = "Backpack Weight to Body Weight Ratio",
       title = "Relationship between Backpack Weight and Body Weight to Backpack Weight Ratio",
       subtitle ="in Males and Females with and without Back Pain (T/F)") +
  theme_bw() +
  theme(plot.title.position = "plot",
         plot.title = element_text(size = 14),
         plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
         axis.text = element_text(size = 8))

# The graph below shows that in both males and females backpack weight to body weight ratio correlates linearly with 
# backpack weight in those with and without back pain.
```
```{r}
backpack %>%
  mutate(back_problems = as.logical(back_problems)) %>%
  ggplot() +
  aes(x = body_weight,  y = ratio) +
  geom_point() +
  geom_smooth(method = "lm", size = 0.5) +
  facet_grid(sex ~ back_problems) +
  labs(x = "Body Weight", y = "Backpack Weight to Body Weight Ratio",
       title = "Relationship between Body Weight and Body Weight to Backpack Weight Ratio",
       subtitle ="in Males and Females with and without Back Pain (True/False)") +
  theme_bw() +
  theme(plot.title.position = "plot",
         plot.title = element_text(size = 14),
         plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
         axis.text = element_text(size = 8))

# The graph below shows that there is no linear relationship between backpack to 
# body weight ratio and body weight in females with no backpain.
# There is a weak inverse linear realtionship backpack to body weight ratio and body weight in males with no backpain.
# There is a weak inverse linear relationship backpack to body weight ratio and body weight in females with backpain.
# There is an inverse linear relationship backpack to body weight ratio and body weight in males with backpain
```
```{r}
backpack %>%
  mutate(back_problems = as.logical(back_problems)) %>%
  ggplot() +
  aes(x = backpack_weight,  y = body_weight) +
  geom_point() +
  geom_smooth(method = "lm", size = 0.5) +
  facet_grid(sex ~ back_problems) +
  labs(x = "Backpack Weight", y = "Body Weight",
       title = "Relationship between Backpack Weight and Body Weight",
       subtitle ="in Males and Females with and without Back Pain (T/F)") +
  theme_bw() +
  theme(plot.title.position = "plot",
         plot.title = element_text(size = 14),
         plot.subtitle = element_text(size = 12),
        axis.title = element_text(size = 10),
         axis.text = element_text(size = 8))

# The graph below shows that there is a weak linear relationship between 
# body weight  and backpack weight in females with no back pain.
# There is a weak  linear relationship backpack to body weight and backpack weight in males with no back pain.
# There is a weak linear relationship backpack to body weight and backpack weight in females with back pain.
# There is an inverse linear relationship in backpack to body weight and backpack weight in males with back pain
```






