---
title: "Hypothesis Test and Probability Homework"
output: html_notebook
---


```{r}

library(tidyverse)
library(infer)

msleep <- ggplot2::msleep

```

Question 1:

```{r}

glimpse(msleep)

```
Question 2:

```{r}

#one sample test - compare mean from one sample against specific value

# H0: sample_mean_sleep_total - jabberwocky_mean_sleep_total == 0
# Ha: sample_mean_sleep_total - jabberwocky_mean_sleep_total != 0
# alpha = 0.05

msleep %>% 
  ggplot(aes(x = sleep_total)) + 
  geom_histogram(bins = 25)


obs_stat <- msleep %>%
  summarise(mean_sleep_total = mean(sleep_total))


null_dist <- msleep %>% 
specify(response = sleep_total) %>% 
hypothesize(null ="point", mu = 7) %>%   
generate(reps = 10000, type = "bootstrap") %>%    
calculate(stat = "mean")



null_dist %>% 
  visualise(bins = 30) +
  shade_p_value(obs_stat = obs_stat$mean_sleep_total, direction = "both")


null_dist %>% 
  get_p_value(obs_stat = obs_stat$mean_sleep_total, 
               direction = "both")

# p-value is very close to 0 and les than 0.05 so based on this and the visualisation of our obs stat on our null distribution we can say that it is unlikley that the mean sleep_total is the same as the typical jabberwocky value.
#Therefore we reject our H0 in favour of Ha.

```

Question 3:

```{r}

# alpha == 0.05
# 
# H0: mean_sleep_total_omni - mean_sleep_total_carni == 0
# 
# Ha:  mean_sleep_total_omni - mean_sleep_total_carni > 0

msleep %>% 
  filter(vore == "omni" | vore == "herbi") %>% 
  ggplot(aes(y = sleep_total, x = vore)) +
  geom_boxplot()


omni_herbi_obs_stat <- msleep %>% 
  filter(vore == "omni" | vore == "herbi") %>% 
  specify(response = sleep_total, explanatory = vore) %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi"))

omni_herbi_null_dist <- msleep %>% 
  filter(vore == "omni" | vore == "herbi") %>% 
  specify(response = sleep_total, explanatory = vore) %>% 
  hypothesise(null = "independence") %>% 
  generate(reps = 10000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("omni", "herbi"))

omni_herbi_null_dist %>% 
  visualise(bins = 25) +
  shade_p_value(omni_herbi_obs_stat$stat, "greater")


omni_herbi_null_dist %>% 
  get_p_value(omni_herbi_obs_stat$stat, "greater") %>% 
  pull() %>% 
  format.pval()

## p value of 0.1191 is greater than 0.05 so we can accept our null hypothesis than omnivores do not sleep significantly longer than herbivores.







```

Question 4:


```{r}

# alpha = 0.05 
# H0 : prop_animals_domesticated == 0.05
# Ha: prop_animals_domesticated > 0.05

observed_stat_ani_dom <- msleep %>% 
  drop_na(conservation) %>% 
  summarise(prop = sum(conservation %in% "domesticated") / n())


msleep_domesticated_prop <- msleep %>% 
  drop_na(conservation) %>% 
  mutate(is_domesticated = ifelse(conservation == "domesticated", "domesticated", "not domesticated")) 

null_distribution <- msleep_domesticated_prop %>% 
  specify(response = is_domesticated,
          success = "domesticated") %>% 
  hypothesise(null = "point", p = 0.05) %>% 
  generate(reps = 2000, type = "draw") %>%  
  calculate(stat = "prop") 



null_dist_dom_ani <- msleep_domesticated_prop %>% 
  specify(response = is_domesticated,
          success = "domesticated") %>% 
  hypothesise(null = "point", p = 0.05) %>% 
  generate(reps = 10000, type = "draw") %>%  
  calculate(stat = "prop") 


null_dist_dom_ani %>% 
  visualise(bins = 15) +
  shade_p_value(observed_stat_ani_dom$prop, "greater")

null_dist_dom_ani %>% 
  get_p_value(observed_stat_ani_dom$prop, "greater")

# p value is 3x10^4 so we can reject the null hypothesis and say that the proportion of domesticated animals is grater than 5%.


```

## Hypothesis Testing 

Question 1:
one-sample test for proportions
H0: prop_pop_coffee_yes = 0.4
Ha: prop_pop_coffee_yes > 0.4
method = draw

Question 2:
paired_two_sample_test
H0: diff_in_propn_users_clicked_banner_b - propn_users_clicked_banner_a = 0
Ha: diff_in_propn_users_clicked_banner_b - propn_users_clicked_banner_a > 0
method = permute

Question 3:
one sample t test
H0: mean_parts_width = 145
Ha: mean_parts_width != 145



Question 1.

Coffee shop problem. Significance level: 0.05, calculated p
-value: 0.07

p value is greater than 0.05 so we accept null hypothesis than prop_pop_coffee = 0.4


Question 2.

Website company problem. Significance level: 0.01, p
-value: 0.006

p value is less than 0.01 so we reject the null hypothesis and suggest our ha is correct that Ha: diff_in_propn_users_clicked_banner_b - propn_users_clicked_banner_a > 0

Question 3.

Manufacturing company problem. Significance level: 0.05, p
-value: 0.55
p value greatr than 0.05 so we can accept the null hypothesis that nea_parts_width = 145


Extension:

```{r}
library(tidyverse)
transactions <- read_csv("online_retail_subset.csv")
```

Question 1.

sup(A)= number of transactions involving A / total number of transactions

```{r}

sup_a <- transactions %>% 
  summarise(propn = sum(StockCode == 22469)/n())
  

```


Question 2:

sup(A→B)=P(A and B being purchased together) = number of transactions involving A and B / total number of transactions

```{r}

stock_code_21110 <- transactions %>% 
  filter(StockCode == 21110)

stock_code_22469 <- transactions %>% 
  filter(StockCode == 22469)

# a_b_transactions <- inner_join(stock_code_22469, stock_code_21110, "InvoiceNo")

a_b_transactions_2 <- inner_join(stock_code_22469, stock_code_21110, "InvoiceDate") %>% 
  nrow()

sup_a_b <- transactions %>% 
  summarise(propn = sum(a_b_transactions_2)/n())

```
conf(A→B)=P(A and B being purchased together)P(A being purchased)

```{r}

conf_a_b <- sup_a_b / sup_a

```



Question 3:

lift(A→B)=sup(A→B)sup(A)×sup(B)

```{r}
sup_b <- transactions %>% 
  summarise(propn = sum(StockCode == 21110)/n())


lift_a_b <- sup_a_b / (sup_a * sup_b)
```


Apriori algorithm

```{r}
library(arules)
library(arulesViz)
```

```{r}
transactions_reformat <- transactions %>%
  select(InvoiceNo, Description) %>%
  na.omit()

write_csv(transactions_reformat, "transactions_reformat.csv")

apriori_format <- read.transactions("transactions_reformat.csv", format = "single", sep = ",", header = TRUE, cols = c("InvoiceNo", "Description"))
```
```{r}
inspect(head(apriori_format))
```

